---
title: "GM_loop"
author: "ChenFengling"
date: "2018Äê6ÔÂ18ÈÕ"
output: html_document
---
**loop ananylis of C2C12**

## Use HiCloop to call loops in 10Kb and adjust parameters

### select parameter
```{bash, eval=FALSE}
# dir /home/fchen/data/zhudahai_insitu_process
# awk '{if($7<0.001 && $8>0.7) print }' DM_cnn_loop.bedpe |wc -l
24051
# awk '{if($7<0.001 && $8>0.8) print }' DM_cnn_loop.bedpe |wc -l
23147
# awk '{if($7<0.001 && $8>0.8) print }' GM_cnn_loop.bedpe |wc -l
23433
# awk '{if($7<0.001 && $8>0.9) print }' GM_cnn_loop.bedpe |wc -l
21840
# awk '{if($7<0.001 && $8>0.9) print }' DM_cnn_loop.bedpe |wc -l
21530
# awk '{if($7<0.0001 && $8>0.9) print }' DM_cnn_loop.bedpe |wc -l
19813
# awk '{if($7<0.0001 && $8>0.9) print }' GM_cnn_loop.bedpe |wc -l
19064
# awk '{if($7<0.00001 && $8>0.9) print }' DM_cnn_loop.bedpe |wc -l
18156
# awk '{if($7<0.00001 && $8>0.9) print }' GM_cnn_loop.bedpe |wc -l
16870

name=GM
awk '{if($7<0.0001 && $8>0.9) print }' ${name}_cnn_loop.bedpe >${name}_loop.bedpe
wc -l ${name}_loop.bedpe
awk -v OFS="\t" '{print $1,$2-10000,$3+10000,$4,$5-10000,$6+10000}' ${name}_loop.bedpe >${name}_loop_extend.bedpe
name=DM
awk '{if($7<0.0001 && $8>0.9) print }' ${name}_cnn_loop.bedpe >${name}_loop.bedpe
wc -l ${name}_loop.bedpe
awk -v OFS="\t" '{print $1,$2-10000,$3+10000,$4,$5-10000,$6+10000}' ${name}_loop.bedpe >${name}_loop_extend.bedpe
```
### CTCF motif enrichment 
```{bash, eval=FALSE,echo=FALSE}
JUICER=/home/software/juicer/CPU
sed 's/chr//g'  GM_loop.bedpe > GM_loop.bedpe1
sed 's/chr//g'  DM_loop.bedpe > DM_loop.bedpe1

time -p java -jar ${JUICER}/juicer_tools.jar motifs mm9 /home/fchen/data/zhudahai_insitu_process/GM_CTCF GM_loop.bedpe1 mm9.motifs.txt
time -p java -jar ${JUICER}/juicer_tools.jar motifs mm9 /home/fchen/data/zhudahai_insitu_process/DM_CTCF DM_loop.bedpe1 mm9.motifs.txt
cut -f16,21 DM_loop.bedpe1_with_motifs.bedpe |sort -k1,1 -k2,2 |uniq -c >DM_motif_orientation.txt
cut -f16,21 GM_loop.bedpe1_with_motifs.bedpe |sort -k1,1 -k2,2 |uniq -c >GM_motif_orientation.txt

```
```{r, echo=FALSE,message=FALSE,results=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
```

```{r }
GM=read.table("GM_motif_orientation.txt")
DM=read.table("DM_motif_orientation.txt")
data_ori <- "GM;DM
--;565;629
++;544;618
-+;56;56
+-;3587;3992"

data <- read.table(text=data_ori, header=T, row.names=1, sep=";", quote="")
data
data_rownames <- rownames(data)
data_colnames <- colnames(data)
data$gene <- data_rownames
data_m <- melt(data, id.vars=c("gene"))
data_m <- data_m %>% group_by(variable) %>% mutate(count=sum(value)) %>% mutate(freq=round(100*value/count,2))
p <- ggplot(data_m, aes(x=variable, y=value, group=gene)) +
	geom_bar(stat="identity", position="fill", aes(fill=gene)) +
	scale_y_continuous(labels = scales::percent) +
	geom_text(aes(label=freq), position=position_fill(vjust=0.5))+theme_set(theme_bw())+theme(panel.grid.major=element_line(colour=NA))
plot(p)
pdf("CTCF_motif_orientation.pdf",height=4,width = 3.5)
plot(p)
dev.off()
```

see 'CTCF_motif_orientation.pdf'

[bar plot in R](http://blog.genesino.com/2017/08/barplot-r/)

## differential loops
### differential from intersection
```{bash, eval=FALSE}
pairToPair -type notboth -is -a GM_loop_extend.bedpe -b DM_loop.bedpe >lost_loop_extend.bedpe
wc -l lost_loop_extend.bedpe
pairToPair -type notboth -is -a DM_loop_extend.bedpe -b GM_loop.bedpe >gained_loop_extend.bedpe
wc -l gained_loop_extend.bedpe
```
19064 GM_loop.bedpe
19813 DM_loop.bedpe
4246 lost_loop_extend.bedpe
4995 gained_loop_extend.bedpe

###  differential loops from Phanstiel et al.

#### step 1. prepare the data

```{bash eval=FALSE}
# dir /home/fchen/HiCpipe/pipe/loop 
nohup time matlab -r "addpath(genpath('/home/fchen/HiCpipe/pipe/loop/'));hicfile={'/home/fchen/data/zhudahai_insitu_process/GM1/maps/'};\
loopfile='/home/fchen/data/zhudahai_insitu_process/all_loops.bedpe';\
name='GM1';resolution=10000;extractpixels(hicfile,loopfile,name,resolution);exit;" > GM1.log </dev/null &

```
#### step 2. call Dloop
allpixels.txt  normfactor.txt
```{r diffloop}
setwd("D:/HiCpipe/zhudahai/C2C12/loop")

library("BiocParallel")
register(MulticoreParam(20))
library("DESeq2")
coldata=data.frame(sample=c(rep("GM",97*3),rep("DM",97*3)),rep=factor(rep(rep(1:3,each=97),2)),pixel_type=rep(c("L",rep("B",96)),6))
rownames(coldata)=c(paste("GM",1:(97*3),sep="_"),paste("DM",1:(97*3),sep="_"))
cts=read.table("allpixels.txt",stringsAsFactors = FALSE,header = FALSE)
loop=cts[,1:3]
cts=cts[,-c(1:3,1:3+100,1:3+200,1:3+300,1:3+400,1:3+500)]
colnames(cts)=rownames(coldata)
# dds <- DESeqDataSetFromMatrix(countData = cts,
#                               colData = coldata,
#                               design = ~ rep+rep:sample+pixel_type+sample:pixel_type+sample)
normFactors=as.matrix(read.table("normfactor.txt",stringsAsFactors = FALSE,header = FALSE))
normFactors=normFactors[,-c(1:3,1:3+100,1:3+200,1:3+300,1:3+400,1:3+500)]
normFactors <- normFactors / exp(rowMeans(log(normFactors)))
colnames(normFactors)=rownames(coldata)

dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = coldata,
                              design = ~ sample+sample:rep+sample:pixel_type)

normalizationFactors(dds) <- normFactors

dds <- DESeq(dds,betaPrior=FALSE,fitType ="local",parallel=TRUE)
# find count of loops are differential compared to the background in DM vs GM :
res=results(dds, contrast=list("sampleGM.pixel_typeL","sampleDM.pixel_typeL"))
  

res <- results(dds, contrast=c("sample","GM","DM"),alpha=0.001 )
res <- as.data.frame(res)
summary(res)
plotMA(res, ylim=c(-3,3))

```

annotate the loops for genes and distial
```{bash , eval=FALSE}
name=lost_loop
mkdir tmp
bedtools intersect -a ${name}_extend.bedpe -b /home/reference/mouse/mm9/tss.bed -wao|sort -k1,1 -k2,2n -k4,4 -k5,5n -u >tmp/${name}_1
awk -v OFS="\t" '{print $4,$5,$6,$1,$2,$3,$11}' tmp/${name}_1>tmp/${name}_2
bedtools intersect -a tmp/${name}_2  -b /home/reference/mouse/mm9/tss.bed -wao|sort -k1,1 -k2,2n -k4,4 -k5,5n -u >tmp/${name}_3
awk -v OFS="\t" '{print $4,$5,$6,$1,$2,$3,$7,$12}' tmp/${name}_3>tmp/${name}.anno.bedpe
awk -v OFS="\t" '{if($7=="-1" && $8!="-1") print $4,$5,$6,$1,$2,$3,$8,$7; else print $0}' tmp/${name}.anno.bedpe>${name}.anno.bedpe
awk -v OFS="\t" '{print $1,$2,$3,$7,".","."}' ${name}.anno.bedpe>${name}_anchor1.bed
awk -v OFS="\t" '{print $4,$5,$6,$8,".","."}' ${name}.anno.bedpe>${name}_anchor2.bed
```
